---
title: "Alpha diversity analysis fungi"
author: "Valentin Gfeller"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#global knitter option
knitr::opts_chunk$set(fig.align = "center", echo = FALSE)

# load packages 
library(tidyverse); packageVersion("tidyverse")
library(phyloseq); packageVersion("phyloseq")
library(vegan); packageVersion("vegan")
library(ape); packageVersion("ape")
library(here); packageVersion("here")
library(readxl); packageVersion("readxl")
library(ggpubr); packageVersion("ggpubr")
library(nlme); packageVersion("nlme")
library(lme4); packageVersion("lme4")
library(car); packageVersion("car")
library(ggbeeswarm); packageVersion("ggbeeswarm")
library(emmeans); packageVersion("emmeans")
library(ggtext); packageVersion("ggtext")
library(cowplot); packageVersion("cowplot")


# functions were written for repeatedly used code (see "/Scripts/Functions.R")
# source functions
source(here("Scripts", "Functions.R"))

# set global option for ggplot2
theme_set(theme_bw(base_size = 12))
```


# Import data and rarefy data
```{r include=FALSE}
load(here("Data", "Interim", "phy.fun.all.RData"))

# Rarefy
phy.fun_rar <- rarefy_even_depth(phy.fun.all, rngseed = 5070, sample.size = 1000)


```


# Calculate diversity indices
## Richness
((rarefying --> richness) x 1000) --> mean
```{r}
# Get OTU table
d.samp <- phy.fun.all %>% sample_data() %>% data.frame()
d.otu <- phy.fun.all %>% otu_table() %>% data.frame()

# Get threshold to rarefy
thresh <- sort(colSums(d.otu), decr = FALSE)[1:10]
thresh <- floor(thresh[1]/ 100) * 100 # round numbers to the next higher 100

# Function to calculate species richness by rarefying and calculating the richness 1000 times
calc_sp_richness <- function(d.otu, thresh){
  # Calculate species richness
  t.sp_richness <- c()
  for (i in 1:1000){
    # rarefy OTU table
    temp_sp_richness <- t(rrarefy(t(d.otu), thresh)) %>% as.matrix()
    sp_richness <- vegan::specnumber(temp_sp_richness, MARGIN = 2) 
    t.sp_richness <- cbind(t.sp_richness, sp_richness)
    }
    return(t.sp_richness)
}

# Calculate Shannon index
set.seed(5070)
x1000_sp_richness <- calc_sp_richness(d.otu, thresh)
x1000_sp_richness[1:5, 1:5]
d.alpha <- bind_cols(d.samp, sp_richness = rowMeans(x1000_sp_richness) %>% 
      round()) %>% 
  select(1:13, "richness" = sp_richness, everything())
d.alpha[1:5, 10:14]

```

## Shannon Index
((rarefying --> Shannon Index) x 1000) --> mean
```{r}
# Function to calculate Shannon Index by rarefying and calculating the Shannon Index 1000 times
calc_shannon <- function(d.otu, thresh){
  # Calculate Shannon diversity
  t.shannon <- c()
  for (i in 1:1000){
    # rarefy OTU table
    temp_shannon <- t(rrarefy(t(d.otu), thresh)) %>% as.matrix()
    shannon <- vegan::diversity(temp_shannon, index = "shannon", MARGIN = 2)
    t.shannon <- cbind(t.shannon, shannon)
    }
    return(t.shannon)
}

# Calculate Shannon index
set.seed(5070)
x1000_Shannon <- calc_shannon(d.otu, thresh)
x1000_Shannon[1:5, 1:5]
d.alpha <- bind_cols(d.alpha, d = rowMeans(x1000_Shannon)) %>% 
  select(1:14, "shannon" = d, everything())
d.alpha[1:5, 10:15]

rm(list = c("x1000_Shannon", "x1000_sp_richness"))
```

## Pielou's evenness
Pielou's evenness: shannon / log(richness)
```{r}
# Calculate Pielou's evenness
d.alpha <- d.alpha %>% 
  mutate(pielou = shannon/log(richness)) %>% 
  select(1:15, pielou, everything())
d.alpha[1:5, 10:16]

cor(d.alpha$shannon, d.alpha$pielou)
plot(d.alpha$shannon, d.alpha$pielou) # Shannon and Pielou are highly correlated! --> only analyze Shannon

```

Save alpha diversity measures
```{r}
t.comp_alpha_div <- d.alpha %>% 
  select("pea_line" = genotype, "Replicate" = rep, "Richness" = richness,
         "Shannon" = shannon) %>% 
  mutate(letter = str_replace_all(pea_line, "[:digit:]", ""),
         number = parse_number(pea_line),
         pea_line2 = paste0(letter, sprintf("%03.0f", number))) %>% 
  select(-c(letter, number)) %>% 
  select(pea_line, pea_line2, everything()) %>% 
  arrange(pea_line2) %>% tibble()

write.csv(t.comp_alpha_div, row.names = FALSE,
          file = here("Data", "Interim", "alpha_div_fun.csv"))
```

# Statistics and visualization
# Shannon
## Seed source
```{r}
# Breeding material effect
mod.mat <- lm(shannon ~ breeding_material, data = d.alpha)
Anova(mod.mat)
summary(mod.mat)
plot.mod.vg(mod.mat)

lab_anova_source <- paste0("**ANOVA**<Br> Seed source: *R<sup>2</sup>* = ", 
                           signif(summary(mod.mat)$adj.r.squared*100, 2),"%;  *P* = ",
                           signif(Anova(mod.mat)$`Pr(>F)`[1], 2))


## Pairwise comparison
emmeans(mod.mat, specs = pairwise ~ breeding_material) %>% 
  .$contrasts %>%
     rbind(adjust = "fdr") %>% 
     summary(infer = TRUE)

# calculate differences
t.mult_cld <- multcomp::cld(emmeans(mod.mat, specs = ~ breeding_material), 
                            Letters = letters, adjust = "fdr") %>% as.data.frame()

t.pos_cld_ <- d.alpha %>%
  group_by(breeding_material) %>% 
  summarise(pos_cld = max(shannon) + 0.25)

t.cld <- left_join(t.mult_cld, t.pos_cld_)

```

```{r}
## Plot
p.bred <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>% 
  ggplot(aes(x = breeding_material, 
             y = shannon))  +
  geom_boxplot(outlier.colour = NA) +
  geom_quasirandom(aes(color = breeding_material),
                   dodge.width = 0.75, width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 16) +
  stat_summary(fun = mean, 
               geom = "point",
               size = 2,
               shape = 18) +
  stat_summary(fun.data = mean_se, 
               fun.args = list(mult = 1),
               geom = "errorbar",
               width = 0.3)  +
  labs(x = "Seed source",
       y = "Shannon") +
  scale_x_discrete(labels = c(gb = "Gene bank",
                              bm ="Breeding material",
                              cv = "Cultivar")) +
  scale_color_manual(values = c(gb = "#FFCC22", 
                                bm = "#56B4E9", 
                                cv = "#009E73")) 


p <- p.bred +
      geom_text(data = t.cld, aes(x = breeding_material, y = pos_cld, 
                              label = trimws(.group)), size = 5) +
      geom_richtext(aes(x = -Inf, y = 4, label = lab_anova_source),
                    stat = "unique",
                    fill = "white", label.color = "white",
                    label.padding = grid::unit(c(0.5, 0.5, 0.5, 0.75), "lines"),
                    hjust = 0, vjust = 0.5, size = 3,
                    label.r = unit(0, "lines")) +
  ggtitle("Fungi") +
  theme(plot.title = element_text(hjust = 0.05))

p

```

## Genotype
```{r}
# Genotype effect
mod.gen <- lm(shannon ~ genotype, data = d.alpha)
Anova(mod.gen)
summary(mod.gen)$adj.r.square
summary(mod.gen)$r.square
plot.mod.vg(mod.gen)

# lab_anova_gen <- paste0("**ANOVA**<Br> Genotype:  *P* = ", signif(Anova(mod.gen)$`Pr(>F)`[1], 2))
lab_anova_gen <- paste0("**ANOVA**<Br> Genotype: *R<sup>2</sup>* = ", signif(summary(mod.gen)$adj.r.squared*100, 2),"%;  *P* = ", signif(Anova(mod.gen)$`Pr(>F)`[1], 2))

## Plot
v.sort <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>%
  group_by(genotype) %>% 
  summarise(
    mean_shannon = mean(shannon, na.rm = TRUE)) %>% 
  arrange(mean_shannon) %>% 
  pull(genotype) %>% unique()

p.lines <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>%
  mutate(genotype = factor(genotype, levels = v.sort))  %>% 
  ggplot(aes(x = genotype, 
             y = shannon, color = breeding_material)) +
  geom_point(alpha = 0.8, size = 0.5) +
  stat_summary(fun = mean, 
               geom = "point",
               alpha = 0.5) +
  stat_summary(fun.data = mean_se, 
               fun.args = list(mult = 1),
               geom = "errorbar",
               linewidth  = 0.3,
               width = 1.25) +
  scale_color_manual(name = "Seed source",
                     values = c(gb = "#FFCC22", 
                                bm = "#56B4E9", 
                                cv = "#009E73"),
                     labels = c(gb = "Gene bank",
                                bm = "Breeding material",
                                cv = "Cultivar")) +
  labs(y = "Shannon",
       x = "Genotypes")  +
  theme(panel.grid.major = element_blank(), 
        panel.grid.major.y = element_line(linewidth = .1),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 1.75),
        legend.position = "top")
p.lines
p.lines <- p.lines +
      geom_richtext(aes(x = -Inf, y = 4, label = lab_anova_gen, color = NULL),
                    stat = "unique",
                    fill = "white", label.color = "white",
                    label.padding = grid::unit(c(0.5, 0.5, 0.5, 0.75), "lines"),
                    hjust = 0, vjust = 0.5, size = 3,
                    label.r = unit(0, "lines"), show.legend = FALSE)  +
  ggtitle("Fungi") +
  guides(color = guide_legend(direction = "horizontal")) +
  theme(plot.title = element_text(size = 16, hjust = 0.01),
        axis.title = element_text(size = 14),
        legend.position = c(1, 1.05), 
        legend.justification = "right",
        legend.background = element_blank()) 

p.lines

```

## Combine
```{r}
fig <- plot_grid(
               p.lines,
               p + theme(legend.position="none"),
               labels = c("(a)", "(b)"),
               align = "h",
               nrow = 1, rel_widths = c(1, 0.6))

fig

p <- fig
saveRDS(p, here("Data", "Interim", "Figures", "Shannon_both_fungi_stats.rds"))

rm(list = c("p", "fig"))
```


# Richness
## Seed source
```{r}
# Breeding material effect
mod.mat <- lm(richness ~ breeding_material, data = d.alpha)
Anova(mod.mat)
summary(mod.mat)
plot.mod.vg(mod.mat)

lab_anova_source <- paste0("**ANOVA**<Br> Seed source: *R<sup>2</sup>* = ", signif(summary(mod.mat)$adj.r.squared*100, 2),"%;  *P* = ", signif(Anova(mod.mat)$`Pr(>F)`[1], 2))


## Pairwise comparison
emmeans(mod.mat, specs = pairwise ~ breeding_material) %>% 
  .$contrasts %>%
     rbind(adjust = "fdr") %>% 
     summary(infer = TRUE)

# calculate differences
t.mult_cld <- multcomp::cld(emmeans(mod.mat, specs = ~ breeding_material), 
                            Letters = letters, adjust = "fdr") %>% as.data.frame()

t.pos_cld_ <- d.alpha %>%
  group_by(breeding_material) %>% 
  summarise(pos_cld = max(richness) + 5)

t.cld <- left_join(t.mult_cld, t.pos_cld_)

```

```{r}
## Richness
p.bred <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>% 
  ggplot(aes(x = breeding_material, 
             y = richness))  +
  geom_boxplot(outlier.colour = NA) +
  geom_quasirandom(aes(color = breeding_material),
                   dodge.width = 0.75, width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 16) +
  stat_summary(fun = mean, 
               geom = "point",
               size = 2,
               shape = 18) +
  stat_summary(fun.data = mean_se, 
               fun.args = list(mult = 1),
               geom = "errorbar",
               width = 0.3)  +
  labs(x = "Seed source",
       y = "OTU Richness") +
  scale_x_discrete(labels = c(gb = "Gene bank",
                              bm ="Breeding material",
                              cv = "Cultivar")) +
  scale_color_manual(values = c(gb = "#FFCC22", 
                                bm = "#56B4E9", 
                                cv = "#009E73")) 


p <- p.bred +
      geom_text(data = t.cld, aes(x = breeding_material, y = pos_cld, 
                              label = trimws(.group)), size = 5) +
      geom_richtext(aes(x = -Inf, y = 112.5, label = lab_anova_source),
                    stat = "unique",
                    fill = "white", label.color = "white",
                    label.padding = grid::unit(c(0.5, 0.5, 0.5, 0.75), "lines"),
                    hjust = 0, vjust = 0.5, size = 3,
                    label.r = unit(0, "lines")) +  
  ggtitle("Fungi") +
  theme(plot.title = element_text(hjust = 0.05))

p

```


## Genotype
```{r}
# Genotype effect
mod.gen <- lm(richness ~ genotype, data = d.alpha)
Anova(mod.gen)
summary(mod.gen)$adj.r.square
summary(mod.gen)$r.square
plot.mod.vg(mod.gen)

# lab_anova_gen <- paste0("**ANOVA**<Br> Genotype:  *P* = ", signif(Anova(mod.gen)$`Pr(>F)`[1], 2))
lab_anova_gen <- paste0("**ANOVA**<Br> Genotype: *R<sup>2</sup>* = ", signif(summary(mod.gen)$adj.r.squared*100, 2),"%;  *P* = ", signif(Anova(mod.gen)$`Pr(>F)`[1], 2))


## Richness
v.sort <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>%
  group_by(genotype) %>% 
  summarise(
    mean_richness = mean(richness, na.rm = TRUE)) %>% 
  arrange(mean_richness) %>% 
  pull(genotype) %>% unique()

p.lines <- d.alpha %>% 
  mutate(breeding_material = factor(breeding_material, 
                                    levels = c("gb", "bm", "cv"))) %>%
  mutate(genotype = factor(genotype, levels = v.sort))  %>% 
  ggplot(aes(x = genotype, 
             y = richness, color = breeding_material)) +
  geom_point(alpha = 0.8, size = 0.5) +
  stat_summary(fun = mean, 
               geom = "point",
               alpha = 0.5) +
  stat_summary(fun.data = mean_se, 
               fun.args = list(mult = 1),
               geom = "errorbar",
               linewidth  = 0.3,
               width = 1.25) +
  scale_color_manual(name = "Seed source",
                     values = c(gb = "#FFCC22", 
                                bm = "#56B4E9", 
                                cv = "#009E73"),
                     labels = c(gb = "Gene bank",
                                bm = "Breeding material",
                                cv = "Cultivar")) +
  labs(y = "OTU Richness",
       x = "Genotypes")  +
  theme(panel.grid.major = element_blank(), 
        panel.grid.major.y = element_line(linewidth = .1),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 1.75),
        legend.position = "top")
p.lines
p.lines <- p.lines +
      geom_richtext(aes(x = -Inf, y = 110, label = lab_anova_gen, color = NULL),
                    stat = "unique",
                    fill = "white", label.color = "white",
                    label.padding = grid::unit(c(0.5, 0.5, 0.5, 0.75), "lines"),
                    hjust = 0, vjust = 0.5, size = 3,
                    label.r = unit(0, "lines"), show.legend = FALSE) +
  ggtitle("Fungi") +
  guides(color = guide_legend(direction = "horizontal")) +
  theme(plot.title = element_text(size = 16, hjust = 0.01),
        axis.title = element_text(size = 14),
        legend.position = c(1, 1.05), 
        legend.justification = "right",
        legend.background = element_blank()) 


```


## Combine
```{r}
fig <- plot_grid(
               p.lines,
               p + theme(legend.position = "none"),
               labels = c("(a)", "(b)"),
               align = "h",
               nrow = 1, rel_widths = c(1, 0.6))

fig

p <- fig
saveRDS(p, here("Data", "Interim", "Figures", "Richness_both_fungi_stats.rds"))

```

# Alpha diversity and plant Resistance
## Emergence
```{r}
# set global option for ggplot2
theme_set(theme_bw(base_size = 14))

# add emergence as factor
d.alpha <- d.alpha %>%
  mutate(emergence_f = factor(emergence))

# Emergence
## Richness
p.alpha_emerg_rich <- d.alpha %>% 
  ggplot(aes(x = emergence, 
             y = richness)) +
  geom_quasirandom(aes(fill = emergence_f),
                   dodge.width = 0.7, width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "Emergence",
       y = "OTU Richness") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "grey",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5)  +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  gg_fill_emerg() +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)

p.alpha_emerg_rich


## Shannon
p.alpha_emerg_shann <- d.alpha %>% 
  ggplot(aes(x = emergence, 
             y = shannon)) +
  geom_quasirandom(aes(fill = emergence_f),
                   dodge.width = 0.7, width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "Emergence",
       y = "Shannon") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "black",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  gg_fill_emerg() +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)


p.alpha_emerg_shann

```

## RRI
```{r}
# RRI
## Richness
d.alpha <- d.alpha %>% 
  mutate(RRI_f = factor(RRI))

fill_rri <- d.alpha %>% 
  mutate(RRI_col = case_when(
    RRI < 3.5   ~ "#5ec962",
    RRI < 4.5   ~ "#21918c",
    RRI < 5.5   ~ "#3b528b",
    TRUE        ~ "#440154"))  %>% pull(RRI_col, RRI_f)

p.alpha_rri_rich <- d.alpha %>%  
  ggplot(aes(x = RRI, 
             y = richness)) +
  geom_quasirandom(aes(fill = RRI_f),
                   width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "RRI",
       y = "OTU Richness") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "black",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  scale_fill_manual(values = fill_rri) +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)

  

p.alpha_rri_rich


## Shannon
p.alpha_rri_shann <- d.alpha %>%  
  ggplot(aes(x = RRI, 
             y = shannon)) +
  geom_quasirandom(aes(fill = RRI_f),
                   width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "RRI",
       y = "Shannon") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "darkgrey",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  scale_fill_manual(values = fill_rri) +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)

  

p.alpha_rri_shann

```


## Combine
```{r}
# Shannon
fig <- plot_grid(
               p.alpha_emerg_shann + ggtitle("Fungi")+
                 theme(legend.position="none",
                       plot.title = element_text(hjust = 0.04)),
               p.alpha_rri_shann + ggtitle("Fungi") +
                 theme(legend.position="none",
                       plot.title = element_text(hjust = 0.04)),
               # labels = c("(a)", "(b)"),
               align = "h",
               nrow = 1, rel_widths = c(1, 1))

fig

# To save the data for the regression
p <- plot_grid(fig)
saveRDS(p, here("Data", "Interim", "Figures", "Shannon_resistance_fungi.rds"))

rm(list = c("p", "fig"))
```


```{r}
# Richness
fig <- plot_grid(
               p.alpha_emerg_rich + ggtitle("Fungi")+
                 theme(legend.position="none",
                       plot.title = element_text(hjust = 0.04)),
               p.alpha_rri_rich + ggtitle("Fungi") +
                 theme(legend.position="none",
                       plot.title = element_text(hjust = 0.04)),
               # labels = c("(a)", "(b)"),
               align = "h",
               nrow = 1, rel_widths = c(1, 1))

fig


# To save the data for the regression
p <- plot_grid(fig)

saveRDS(p, here("Data", "Interim", "Figures", "Richness_resistance_fungi.rds"))

rm(list = c("p", "fig"))

```
# Test if the two datapoints with RRI 2 drive the correlations
## RRI

```{r}
d.alpha <- d.alpha %>% 
  filter(RRI > 2)
```


```{r}
# RRI
## Richness
d.alpha <- d.alpha %>% 
  mutate(RRI_f = factor(RRI))

fill_rri <- d.alpha %>% 
  mutate(RRI_col = case_when(
    RRI < 3.5   ~ "#5ec962",
    RRI < 4.5   ~ "#21918c",
    RRI < 5.5   ~ "#3b528b",
    TRUE        ~ "#440154"))  %>% pull(RRI_col, RRI_f)

p.alpha_rri_rich <- d.alpha %>%  
  ggplot(aes(x = RRI, 
             y = richness)) +
  geom_quasirandom(aes(fill = RRI_f),
                   width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "RRI",
       y = "OTU Richness") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "black",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  scale_fill_manual(values = fill_rri) +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)

  

p.alpha_rri_rich


## Shannon
p.alpha_rri_shann <- d.alpha %>%  
  ggplot(aes(x = RRI, 
             y = shannon)) +
  geom_quasirandom(aes(fill = RRI_f),
                   width = 0.15, alpha = 0.5,
                   size = 2, show.legend = FALSE, shape = 21) +
  labs(x = "RRI",
       y = "Shannon") +
  geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, color = "darkgrey",
              fullrange = TRUE) +
  ggpubr::stat_cor(aes(label = 
                         paste(after_stat(rr.label), after_stat(..p.label..),
                               sep = "~`,`~")), method = "spearman",
                   label.x.npc = "right", hjust = 1.1, vjust = -0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
  scale_fill_manual(values = fill_rri) +
  ggpubr::stat_regline_equation(label.x.npc = "left", vjust = -0.65)

  

p.alpha_rri_shann

```


```{r}

```

