---
title: "Network analysis and hub detection"
author: "Valentin Gfeller"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#global knitter option
knitr::opts_chunk$set(fig.align = "center", echo = FALSE)

# load packages
library(tidyverse); packageVersion("tidyverse")
library(ggbeeswarm); packageVersion("ggbeeswarm")
library(emmeans); packageVersion("emmeans")
library(nlme); packageVersion("nlme")
library(car); packageVersion("car")
library(here); packageVersion("here")
library(ggpubr); packageVersion("ggpubr")
library(knitr); packageVersion("knitr")
library(phyloseq); packageVersion("phyloseq")
library(ape); packageVersion("ape")
library(vegan); packageVersion("vegan")
library(geomtextpath); packageVersion("geomtextpath")
library(SpiecEasi)
library(igraph)

# functions were written for repeatedly used code (see "/Scripts/Functions.R")
# source functions
source(here("Scripts", "Functions.R"))

# set global option for ggplot2
theme_set(theme_bw(base_size = 12))

```

# Data
## Import
```{r include=FALSE}
load(here("Data", "Interim", "phy.bac_all.RData"))
phy.bac_all

load(here("Data", "Interim", "phy.fun.all.RData"))
phy.fun.all

```

## Check data
```{r}
# Look at data
phy.bac_all %>% sample_data() %>% pull(name.ngs_long) %>% .[1:20]
phy.fun.all %>% sample_data() %>% pull(name.ngs) %>% .[1:20]

# Adjust names to combine the two
tab.samp_bact <- phy.bac_all %>% sample_data() %>% pull(name.ngs_long) %>% 
  str_replace("_16S", "") %>% 
  str_replace("REP", "Rep")
data.frame(x = sample_data(phy.bac_all)$name.ngs_long, y = tab.samp_bact) %>% slice_sample(n = 20)
sample_data(phy.bac_all)$name.ngs_long <- tab.samp_bact
sample_names(phy.bac_all) <- tab.samp_bact

tab.samp_fung <- phy.fun.all %>% sample_data() %>% pull(name.ngs)
data.frame(x = sample_names(phy.fun.all), y = tab.samp_fung) %>% slice_sample(n = 20)
sample_names(phy.fun.all) <- tab.samp_fung

# Subset samples to the ones present in bacteria and fungi (intersect)
to_keep <- intersect(tab.samp_bact, tab.samp_fung)

phy.bac_all_red <- subset_samples(phy.bac_all, name.ngs_long %in% to_keep)
phy.fun.all_red <-  subset_samples(phy.fun.all,  name.ngs %in% to_keep)

# Check number of lost samples
phy.bac_all %>% sample_data() %>% dim() %>% .[1]
phy.fun.all %>% sample_data() %>% dim() %>% .[1]
phy.bac_all_red %>% sample_data() %>% dim() %>% .[1]

```

## Combine data
```{r}
d.bact_red <- phy.bac_all_red %>% otu_table() %>% data.frame()
d.fun_red  <- phy.fun.all_red  %>% otu_table() %>% data.frame()
all.equal(d.bact_red %>% colnames() %>% sort(), d.fun_red %>% colnames() %>% sort())

d.otu_comb <- bind_rows(d.bact_red, d.fun_red)
m.otu_comb <- d.otu_comb %>% as.matrix()

phy.comb <- phyloseq(otu_table(m.otu_comb, taxa_are_rows = TRUE),
                     sample_data(phy.fun.all_red))

phy.comb_all <- phy.comb

# Counts
save(phy.comb, file = here("Data", "Interim", "phy.comb_all.RData"))
```

## Import - CLR
```{r include=FALSE}
load(here("Data", "Interim", "phy.bac_clr.RData"))
phy.bac_clr

load(here("Data", "Interim", "phy.fun_clr.RData"))
phy.fun_clr

```

## Check data- CLR
```{r}
# Look at data
phy.bac_clr %>% sample_data() %>% pull(name.ngs_long) %>% .[1:20]
phy.fun_clr %>% sample_data() %>% pull(name.ngs) %>% .[1:20]

# Adjust names to combine the two
tab.samp_bact <- phy.bac_clr %>% sample_data() %>% pull(name.ngs_long) %>% 
  str_replace("_16S", "") %>% 
  str_replace("REP", "Rep")
data.frame(x = sample_data(phy.bac_clr)$name.ngs_long, y = tab.samp_bact) %>% slice_sample(n = 20)
sample_data(phy.bac_clr)$name.ngs_long <- tab.samp_bact
sample_names(phy.bac_clr) <- tab.samp_bact

tab.samp_fung <- phy.fun_clr %>% sample_data() %>% pull(name.ngs)
data.frame(x = sample_names(phy.fun_clr), y = tab.samp_fung) %>% slice_sample(n = 20)
sample_names(phy.fun_clr) <- tab.samp_fung

# Subset samples to the ones present in bacteria and fungi (intersect)
to_keep <- intersect(tab.samp_bact, tab.samp_fung)

phy.bac_clr_red <- subset_samples(phy.bac_clr, name.ngs_long %in% to_keep)
phy.fun_clr_red <-  subset_samples(phy.fun_clr,  name.ngs %in% to_keep)

# Check number of lost samples
phy.bac_clr %>% sample_data() %>% dim() %>% .[1]
phy.fun_clr %>% sample_data() %>% dim() %>% .[1]
phy.bac_clr_red %>% sample_data() %>% dim() %>% .[1]


```

## Combine data - CLR
```{r}
d.bact_red <- phy.bac_clr_red %>% otu_table() %>% data.frame()
d.fun_red  <- phy.fun_clr_red  %>% otu_table() %>% data.frame()
all.equal(d.bact_red %>% colnames() %>% sort(), d.fun_red %>% colnames() %>% sort())

d.otu_comb_clr <- bind_rows(d.bact_red, d.fun_red)
m.otu_comb_clr <- d.otu_comb_clr %>% as.matrix()

phy.comb_clr <- phyloseq(otu_table(m.otu_comb_clr, taxa_are_rows = TRUE),
                     sample_data(phy.fun_clr_red))

phy.comb_clr_all <- phy.comb_clr

# Counts
save(phy.comb_clr, file = here("Data", "Interim", "phy.comb_clr_all.RData"))
```


# Run SpiecEasi pipeline with both kingdoms
```{r}
# Prepare data
phy.comb_fun <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("f"), phy.comb)
phy.comb_bac <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("b"), phy.comb)

# # Run pipeline
# SpEa_mat_comb <- spiec.easi(list(phy.comb_bac, phy.comb_fun),
#                             method = 'mb',
#                             nlambda = 50,
#                             lambda.min.ratio = 1e-2,
#                             pulsar.params = list(rep.num = 50, seed = 5070,
#                                                  ncores = 40))
# 
# saveRDS(SpEa_mat_comb,
#         here("Data", "Interim", "SpiecEasi_object_combined.rds"))

SpEa_mat_comb <- readRDS(here("Data", "Interim", "SpiecEasi_object_combined.rds"))

# create network
set.seed(5070)
igr_net_comb <- adj2igraph(getRefit(SpEa_mat_comb),
                           vertex.attr = list(name = taxa_names(phy.comb)))

degree(igr_net_comb) %>% sort() %>% .[1:10]
degree(igr_net_comb) %>% sort(decreasing = TRUE) %>% .[1:10]

# remove if less than 1 degree
Isolated <-  which(degree(igr_net_comb)==0) %>% names()
igr_net_comb_red  <-  delete.vertices(igr_net_comb, Isolated)

# calculate layout
am.coord <- layout.fruchterman.reingold(igr_net_comb_red)
plot(igr_net_comb_red, layout = am.coord)

# Set color to differentiate the kingdom
n_bact <- V(igr_net_comb_red)$name %>% str_detect("b") %>% sum()
N_fung <- V(igr_net_comb_red)$name %>% str_detect("f") %>% sum()
dtype <- c(rep("darkgrey", n_bact), rep("white", N_fung))

```

# Set color of text to differential abundant OTUs
```{r}
# Load da data for bacteria (RRI) and fungi (emergence)
corr.test <- readRDS(here("Data", "Interim", "DA_bac_corr_RRI_all.rds"))
# Tidy output
tab.otu <- data.frame(corr.test) %>% 
  mutate(OTU = row.names(corr.test)) %>% 
  select(OTU, everything()) %>% 
  arrange(spearman.erho) # sort taxa by Spearman's rho
da.otus_bac <- subset(tab.otu, spearman.eBH < 0.05) ## significant OTUs at the 0.05 level (after correction)

corr.test <- readRDS(here("Data", "Interim", "DA_fun_corr_emergence_all.rds"))
# Tidy output
tab.otu <- data.frame(corr.test) %>% 
  mutate(OTU = row.names(corr.test)) %>% 
  select(OTU, everything()) %>% 
  arrange(spearman.erho) # sort taxa by Spearman's rho
da.otus_fun <- subset(tab.otu, spearman.eBH < 0.05) ## significant OTUs at the 0.05 level (after correction)

rm(list = c("corr.test", "tab.otu"))

tab.otu_da_all_comb <- c(da.otus_fun %>% pull(OTU), da.otus_bac %>% pull(OTU))

dtype.text <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             OTU %in% tab.otu_da_all_comb ~ "red",
             TRUE ~ "black")) %>% 
  pull(color)


plot(igr_net_comb_red,
     vertex.color = dtype,
     vertex.shape = "circle",
     vertex.frame.color = "black",
     vertex.size = 4,
     vertex.label.cex = 0.3,
     vertex.label.color = adjustcolor(dtype.text, alpha.f = 0.85),
     edge.color = adjustcolor("skyblue1", alpha.f = 0.25),
     layout = layout.fruchterman.reingold)


# Set color of Text to heritable OTUs
tab.otu_bac_herit <- read_csv(file = here("Data", "Interim", "heritability_otu_bac.csv"))
tab.otu_bac_herit %>% pull(heritable) %>% table()

tab.otu_fun_herit <- read_csv(file = here("Data", "Interim", "heritability_otu_fun.csv"))
tab.otu_fun_herit %>% pull(heritable) %>% table()

# Select heritable
tab.otu_bac_herit <- tab.otu_bac_herit %>% 
  filter(heritable ==TRUE) %>% pull(OTU)
tab.otu_fun_herit <- tab.otu_fun_herit %>% 
  filter(heritable ==TRUE) %>% pull(OTU)

tab.otu_herit_all_comb <- c(tab.otu_bac_herit, tab.otu_fun_herit)

dtype.text <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             TRUE ~ "black"
           )) %>% 
  pull(color)

plot(igr_net_comb_red,
     vertex.color = dtype,
     vertex.shape = "circle",
     vertex.frame.color = "black",
     vertex.size = 4,
     vertex.label.cex = 0.3,
     vertex.label.color = adjustcolor(dtype.text, alpha.f = 0.85),
     edge.color = adjustcolor("skyblue1", alpha.f = 0.25),
     layout = layout.fruchterman.reingold)

```

# Define color of differential abundant OTUs
```{r}
# Load da data for bacteria (RRI) and fungi (emergence)
corr.test <- readRDS(here("Data", "Interim", "DA_bac_corr_RRI_all.rds"))
# Tidy output
tab.otu <- data.frame(corr.test) %>% 
  mutate(OTU = row.names(corr.test)) %>% 
  select(OTU, everything()) %>% 
  arrange(spearman.erho) # sort taxa by Spearman's rho
da.otus_bac <- subset(tab.otu, spearman.eBH < 0.05) ## significant OTUs at the 0.05 level (after correction)

corr.test <- readRDS(here("Data", "Interim", "DA_fun_corr_emergence_all.rds"))
# Tidy output
tab.otu <- data.frame(corr.test) %>% 
  mutate(OTU = row.names(corr.test)) %>% 
  select(OTU, everything()) %>% 
  arrange(spearman.erho) # sort taxa by Spearman's rho
da.otus_fun <- subset(tab.otu, spearman.eBH < 0.05) ## significant OTUs at the 0.05 level (after correction)

rm(list = c("corr.test", "tab.otu"))

tab.otu_da_all_comb <- c(da.otus_fun %>% pull(OTU), da.otus_bac %>% pull(OTU))

tab.otu_da_neg <- c(da.otus_fun %>% filter(spearman.erho < 0) %>% pull(OTU),
                    da.otus_bac %>% filter(spearman.erho < 0) %>% pull(OTU))
tab.otu_da_pos <- c(da.otus_fun %>% filter(spearman.erho > 0) %>% pull(OTU),
                    da.otus_bac %>% filter(spearman.erho > 0) %>% pull(OTU))

# select only the OTUs that are significantly da for all phenotypes
da_int_bac <- readRDS(here("Data", "Interim", "DA_bac_OTUs_intersect_all_phenotypes.rds"))
da_int_fun <- readRDS(here("Data", "Interim", "DA_fun_OTUs_intersect_all_phenotypes.rds"))
da_int_both <- c(da_int_bac, da_int_fun)

tab.otu_da_all_comb <- tab.otu_da_all_comb[tab.otu_da_all_comb %in% da_int_both]
tab.otu_da_neg <- tab.otu_da_neg[tab.otu_da_neg %in% da_int_both]
tab.otu_da_pos <- tab.otu_da_pos[tab.otu_da_pos %in% da_int_both]

# Add colour of vertex
## Only put labels for da OTUs
dtype.text <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             OTU %in% tab.otu_da_neg ~ "black",
             OTU %in% tab.otu_da_pos ~ "black",
             TRUE ~ NA_character_)) %>% 
  pull(color)

## Define colors for bacteria, fungi and positive or negative da fungi or bacteria
dtype <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             OTU  %>% str_detect("b")  ~ "darkgrey",
             OTU %>% str_detect("f")   ~ "white",
             OTU %in% tab.otu_da_neg | 
               OTU %>% str_detect("b") ~ "darkred",
             OTU %in% tab.otu_da_pos |
               OTU %>% str_detect("b") ~ "darkgreen",
             OTU %in% tab.otu_da_neg | 
               OTU %>% str_detect("f") ~ "red",
             OTU %in% tab.otu_da_pos |
               OTU %>% str_detect("f") ~ "lightgreen",
             TRUE ~ NA_character_),
         color = 
           case_when(
             OTU %in% tab.otu_da_neg & 
               OTU %>% str_detect("b") ~ "darkred",
             OTU %in% tab.otu_da_pos &
               OTU %>% str_detect("b") ~ "darkgreen",
             OTU %in% tab.otu_da_neg & 
               OTU %>% str_detect("f") ~ "red",
             OTU %in% tab.otu_da_pos &
               OTU %>% str_detect("f") ~ "lightgreen",
             TRUE ~ color)) %>% 
  pull(color)

plot(igr_net_comb_red,
     vertex.color = adjustcolor(dtype, alpha.f = 0.75),
     vertex.shape = "circle",
     vertex.frame.color = adjustcolor("black", alpha.f = 0.75),
     vertex.size = 4,
     vertex.label.cex = 0.3,
     vertex.label.color = adjustcolor(dtype.text, alpha.f = 0.5),
     edge.color = adjustcolor("skyblue1", alpha.f = 0.3),
     layout = layout.fruchterman.reingold)



# Add information of heritable OTUs (circle colour == blue)
tab.otu_bac_herit <- read_csv(file = here("Data", "Interim", "heritability_otu_bac.csv"))
tab.otu_bac_herit %>% pull(heritable) %>% table()

tab.otu_fun_herit <- read_csv(file = here("Data", "Interim", "heritability_otu_fun.csv"))
tab.otu_fun_herit %>% pull(heritable) %>% table()


# Select top 58 heritable
tab.otu_herit_all_comb <- bind_rows(tab.otu_bac_herit, tab.otu_fun_herit)

tab.otu_herit_all_comb <- tab.otu_herit_all_comb %>% 
  slice_max(heritability, n = 58) %>% pull(OTU)

dtype.frame <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             OTU %in% tab.otu_herit_all_comb ~ "darkblue",
             TRUE ~ "black"
           )) %>% 
  pull(color)

width.frame <- tibble(OTU = V(igr_net_comb_red)$name) %>% 
  mutate(color = 
           case_when(
             OTU %in% tab.otu_herit_all_comb ~ 2.5,
             TRUE ~ 1)) %>% 
  pull(color)



plot(igr_net_comb_red,
     vertex.color = adjustcolor(dtype, alpha.f = 0.75),
     vertex.shape = "circle",
     vertex.frame.color = adjustcolor(dtype.frame, alpha.f = 0.6),
     vertex.frame.width = width.frame,
     vertex.size = 4,
     vertex.label.cex = 0.4,
     vertex.label.color = adjustcolor(dtype.text, alpha.f = 0.75),
     edge.color = adjustcolor("skyblue1", alpha.f = 0.3),
     layout = layout.fruchterman.reingold)

```

```{r}
# save image plot in Script 22
save.image(file = here("Data", "Interim", "Image_Network.RData"))

```

# Hub taxa detection
```{r echo=TRUE}
# Calculate network attributes (betweenness and degree)
net_bn <- betweenness(igr_net_comb_red)
d.bn <- tibble(OTU = names(net_bn), betweenness = net_bn)

net_dg <- degree(igr_net_comb_red)
d.dg <- tibble(OTU = names(net_dg), degree = net_dg)

d.both <- left_join(d.bn, d.dg)

# Top 10%
top10_bn <- net_bn %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

top10_dg <- net_dg %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

# Add to data frame
d.hub_both <- d.both %>% 
  mutate(
    top10_bn = case_when(OTU %in% top10_bn ~ TRUE,
                         TRUE ~ FALSE),
    top10_dg = case_when(OTU %in% top10_dg ~ TRUE,
                         TRUE ~ FALSE),
    top10_both = case_when(top10_bn == TRUE & top10_dg == TRUE ~ TRUE,
                           TRUE ~FALSE))
```

## Save hub taxa
```{r echo=TRUE}
# save
write.csv(d.hub_both, row.names = FALSE,
          file = here("Data", "Interim", "Hub_taxa_both_kingdoms.csv"))

```

# Check how network attributes correlate
```{r echo=TRUE}
# Check how network attributes correlate
d.hub_both %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~")))

# Highlight OTUs in the top 10% of both degree and betweenness
d.hub_both %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both %>% filter(top10_both == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)

(d.hub_both %>% filter(top10_both == TRUE) %>% nrow())/(d.hub_both %>% nrow())


# Highlight OTUs in the top 10 % of either degree or betweenness
d.hub_both %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both %>% filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2) 

(d.hub_both %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow())/(d.hub_both %>% nrow())


# Check ranks
d.hub_both %>% 
  ggplot(aes(rank(degree), rank(betweenness))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both %>% 
               mutate(degree = rank(degree),
                      betweenness = rank(betweenness)) %>% 
               filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)


# Number of OTUs that are either or 
d.hub_both %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow()
d.hub_both %>% filter(top10_dg == TRUE) %>% nrow()
d.hub_both %>% filter(top10_bn == TRUE) %>% nrow()


intersect(d.hub_both %>% filter(top10_dg == TRUE),
          d.hub_both %>% filter(top10_bn == TRUE)) %>% nrow()

# both
d.hub_both %>% filter(top10_both == TRUE) %>% nrow()

```
# Network anlalysis with top abundant OTUs
To check if we get similar hub taxa when the same number of taxa are used for fungal and bacterial OTUs.

## Run SpiecEasi pipeline with both kingdoms
```{r}
# Prepare data
phy.comb_fun <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("f"), phy.comb)
phy.comb_bac <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("b"), phy.comb)

ntaxa(phy.comb_fun)
ntaxa(phy.comb_bac) # Select top 250 OTUs each

# Filter fungi
n_sum <- taxa_sums(phy.comb_fun) %>% sort(decreasing = TRUE) %>% .[251]
phy.comb_fun <- prune_taxa(taxa_sums(phy.comb_fun) > n_sum, phy.comb_fun)

# Filter bacteria
n_sum <- taxa_sums(phy.comb_bac) %>% sort(decreasing = TRUE) %>% .[251]
phy.comb_bac <- prune_taxa(taxa_sums(phy.comb_bac) > n_sum, phy.comb_bac)

# Filter combined phyloseq object
keep <- c(taxa_names(phy.comb_bac), taxa_names(phy.comb_fun))
phy.comb_red <- prune_taxa(keep, phy.comb)
phy.comb_red

# # Run pipeline
# SpEa_mat_comb_top <- spiec.easi(list(phy.comb_bac, phy.comb_fun),
#                             method = 'mb',
#                             nlambda = 50,
#                             lambda.min.ratio = 1e-2,
#                             pulsar.params = list(rep.num = 50, seed = 5070,
#                                                  ncores = 40))
# 
# saveRDS(SpEa_mat_comb_top,
#         here("Data", "Interim", "SpiecEasi_object_comb_top.rds"))

SpEa_mat_comb_top <- readRDS(here("Data", "Interim", "SpiecEasi_object_comb_top.rds"))

# create network
set.seed(5070)
igr_net_comb_top <- adj2igraph(getRefit(SpEa_mat_comb_top),
                           vertex.attr = list(name = taxa_names(phy.comb_red)))



degree(igr_net_comb_top) %>% sort() %>% .[1:10]
degree(igr_net_comb_top) %>% sort(decreasing = TRUE) %>% .[1:10]

# remove if less than 1 degree
Isolated <-  which(degree(igr_net_comb_top)==0) %>% names()
igr_net_comb_top_red  <-  delete.vertices(igr_net_comb_top, Isolated)

# calculate layout
am.coord <- layout.fruchterman.reingold(igr_net_comb_top_red)
plot(igr_net_comb_top_red, layout = am.coord)
```

## Hub taxa detection
```{r echo=TRUE}
# Calculate network attributes (betweenness and degree)
net_bn <- betweenness(igr_net_comb_top_red)
d.bn <- tibble(OTU = names(net_bn), betweenness = net_bn)

net_dg <- degree(igr_net_comb_top_red)
d.dg <- tibble(OTU = names(net_dg), degree = net_dg)

d.both <- left_join(d.bn, d.dg)

# Top 10%
top10_bn <- net_bn %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

top10_dg <- net_dg %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

# Add to data frame
d.hub_both_topAbun <- d.both %>% 
  mutate(
    top10_bn = case_when(OTU %in% top10_bn ~ TRUE,
                         TRUE ~ FALSE),
    top10_dg = case_when(OTU %in% top10_dg ~ TRUE,
                         TRUE ~ FALSE),
    top10_both = case_when(top10_bn == TRUE & top10_dg == TRUE ~ TRUE,
                           TRUE ~ FALSE))

d.hub_both_topAbun %>% filter(top10_both == TRUE)
```


## Check how network attributes correlate
```{r echo=TRUE}
# Check how network attributes correlate
d.hub_both_topAbun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~")))

# Highlight OTUs in the top 10% of both degree and betweenness
d.hub_both_topAbun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_topAbun %>% filter(top10_both == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)

(d.hub_both_topAbun %>% filter(top10_both == TRUE) %>% nrow())/(d.hub_both_topAbun %>% nrow())


# Highlight OTUs in the top 10 % of either degree or betweenness
d.hub_both_topAbun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_topAbun %>% filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2) 

(d.hub_both_topAbun %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow())/(d.hub_both_topAbun %>% nrow())


# Check ranks
d.hub_both_topAbun %>% 
  ggplot(aes(rank(degree), rank(betweenness))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_topAbun %>% 
               mutate(degree = rank(degree),
                      betweenness = rank(betweenness)) %>% 
               filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)


# Number of OTUs that are either or 
d.hub_both_topAbun %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow()
d.hub_both_topAbun %>% filter(top10_dg == TRUE) %>% nrow()
d.hub_both_topAbun %>% filter(top10_bn == TRUE) %>% nrow()


intersect(d.hub_both_topAbun %>% filter(top10_dg == TRUE),
          d.hub_both_topAbun %>% filter(top10_bn == TRUE)) %>% nrow()

# both
d.hub_both_topAbun %>% filter(top10_both == TRUE) %>% nrow()

```

## Compare different methods
```{r}
## Degree or betweenness in top 10%
top_hub_comb <- d.hub_both %>% filter(top10_bn == TRUE | top10_dg == TRUE ) %>% pull(OTU)
top_hub_topAbun <- d.hub_both_topAbun %>% filter(top10_bn == TRUE | top10_dg == TRUE ) %>% pull(OTU)

# Number of OTUs
top_hub_comb %>% length()
top_hub_topAbun %>% length()

# Visualize intersections
## all three
l.otu_top_hub <- list(top_hub_comb = top_hub_comb,
                      top_hub_topAbun = top_hub_topAbun)

UpSetR::upset(UpSetR::fromList(l.otu_top_hub), order.by = "freq")

## Combined kingdom analysis vs. analysis with only top abundant OTUs
l.otu_top_hub <- list(top_hub_comb = top_hub_comb,
                      top_hub_topAbun = top_hub_topAbun)

UpSetR::upset(UpSetR::fromList(l.otu_top_hub), order.by = "freq")

## Check intersect
top_hub_intersect <- intersect(top_hub_comb, top_hub_topAbun)

# Check intersection with OTUs of most interest
otu_interest <- c("fOTU860", "fOTU1450", "fOTU1557", "bOTU15", "bOTU5", 
                  "fOTU1598", "bOTU1", "fOTU22", "bOTU4", "fOTU1517", "fOTU1508")

top_hub_intersect <- intersect(top_hub_intersect, otu_interest)
top_hub_intersect
sym_diff(top_hub_intersect, otu_interest)

```



# Individual kingdoms for comparison
To check if we get similar hub taxa with this method

## Fungi
### Run SpiecEasi pipeline fungi
```{r}
# Prepare data
phy.comb_fun <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("f"), phy.comb)

# # Run pipeline
# SpEa_mat_fun <- spiec.easi(phy.comb_fun,
#                            method = 'mb',
#                            nlambda = 50,
#                            lambda.min.ratio = 1e-2,
#                            pulsar.params = list(rep.num = 50, seed = 5070,
#                                                 ncores = 40))
# 
# saveRDS(SpEa_mat_fun,
#         here("Data", "Interim", "SpiecEasi_object_fun.rds"))

SpEa_mat_fun <- readRDS(here("Data", "Interim", "SpiecEasi_object_fun.rds"))

# create network
set.seed(5070)
igr_net_fun <- adj2igraph(getRefit(SpEa_mat_fun),
                           vertex.attr = list(name = taxa_names(phy.comb_fun)))

degree(igr_net_fun) %>% sort() %>% .[1:10]
degree(igr_net_fun) %>% sort(decreasing = TRUE) %>% .[1:10]

# remove if less than 1 degree
Isolated <-  which(degree(igr_net_fun)==0) %>% names()
igr_net_fun_red  <-  delete.vertices(igr_net_fun, Isolated)

# calculate layout
am.coord <- layout.fruchterman.reingold(igr_net_fun_red)
plot(igr_net_fun_red, layout = am.coord)

```

### Hub taxa detection fungi
```{r echo=TRUE}
# Calculate network attributes (betweenness and degree)
net_bn <- betweenness(igr_net_fun_red)
d.bn <- tibble(OTU = names(net_bn), betweenness = net_bn)

net_dg <- degree(igr_net_fun_red)
d.dg <- tibble(OTU = names(net_dg), degree = net_dg)

d.both_fun <- left_join(d.bn, d.dg)

# Top 10%
top10_bn <- net_bn %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

top10_dg <- net_dg %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

# Add to data frame
d.hub_both_fun <- d.both_fun %>% 
  mutate(
    top10_bn = case_when(OTU %in% top10_bn ~ TRUE,
                         TRUE ~ FALSE),
    top10_dg = case_when(OTU %in% top10_dg ~ TRUE,
                         TRUE ~ FALSE),
    top10_both_fun = case_when(top10_bn == TRUE & top10_dg == TRUE ~ TRUE,
                           TRUE ~ FALSE))

# save hub taxa
write.csv(d.hub_both_fun, row.names = FALSE,
          file = here("Data", "Interim", "Hub_taxa_fun.csv"))

```

### Check how network attributes correlate fungi
```{r echo=TRUE}
# Check how network attributes correlate
d.hub_both_fun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~")))

# Highlight OTUs in the top 10% of both_fun degree and betweenness
d.hub_both_fun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_fun %>% filter(top10_both_fun == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)

(d.hub_both_fun %>% filter(top10_both_fun == TRUE) %>% nrow())/(d.hub_both_fun %>% nrow())


# Highlight OTUs in the top 10 % of either degree or betweenness
d.hub_both_fun %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_fun %>% filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2) 

(d.hub_both_fun %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow())/(d.hub_both_fun %>% nrow())


# Check ranks
d.hub_both_fun %>% 
  ggplot(aes(rank(degree), rank(betweenness))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_fun %>% 
               mutate(degree = rank(degree),
                      betweenness = rank(betweenness)) %>% 
               filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)


# Number of OTUs that are either or 
d.hub_both_fun %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow()
d.hub_both_fun %>% filter(top10_dg == TRUE) %>% nrow()
d.hub_both_fun %>% filter(top10_bn == TRUE) %>% nrow()


intersect(d.hub_both_fun %>% filter(top10_dg == TRUE),
          d.hub_both_fun %>% filter(top10_bn == TRUE)) %>% nrow()


d.hub_both_fun %>% filter(top10_both_fun == TRUE) %>% nrow()
d.hub_both_fun %>% filter(top10_bn == TRUE | top10_dg == TRUE) %>% nrow()

```

## Bacteria
### Run SpiecEasi pipeline bacteria
```{r}
# Prepare data
phy.comb_bac <- prune_taxa(phy.comb %>% taxa_names() %>% str_detect("b"), phy.comb)

# # Run pipeline
# SpEa_mat_bac <- spiec.easi(phy.comb_bac,
#                            method = 'mb',
#                            nlambda = 50,
#                            lambda.min.ratio = 1e-2,
#                            pulsar.params = list(rep.num = 50, seed = 5070,
#                                                 ncores = 40))
# 
# saveRDS(SpEa_mat_bac,
#         here("Data", "Interim", "SpiecEasi_object_bac.rds"))

SpEa_mat_bac <- readRDS(here("Data", "Interim", "SpiecEasi_object_bac.rds"))

# create network
set.seed(5070)
igr_net_bac <- adj2igraph(getRefit(SpEa_mat_bac),
                           vertex.attr = list(name = taxa_names(phy.comb_bac)))

degree(igr_net_bac) %>% sort() %>% .[1:10]
degree(igr_net_bac) %>% sort(decreasing = TRUE) %>% .[1:10]

# remove if less than 1 degree
Isolated <-  which(degree(igr_net_bac)==0) %>% names()
igr_net_bac_red  <-  delete.vertices(igr_net_bac, Isolated)

# calculate layout
am.coord <- layout.fruchterman.reingold(igr_net_bac_red)
plot(igr_net_bac_red, layout = am.coord)

```

### Hub taxa detection bacteria
```{r echo=TRUE}
# Calculate network attributes (betweenness and degree)
net_bn <- betweenness(igr_net_bac_red)
d.bn <- tibble(OTU = names(net_bn), betweenness = net_bn)

net_dg <- degree(igr_net_bac_red)
d.dg <- tibble(OTU = names(net_dg), degree = net_dg)

d.both_bac <- left_join(d.bn, d.dg)

# Top 10%
top10_bn <- net_bn %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

top10_dg <- net_dg %>% sort(decreasing = TRUE) %>% 
  head(n = length(.) * 0.1) %>% names()

# Add to data frame
d.hub_both_bac <- d.both_bac %>% 
  mutate(
    top10_bn = case_when(OTU %in% top10_bn ~ TRUE,
                         TRUE ~ FALSE),
    top10_dg = case_when(OTU %in% top10_dg ~ TRUE,
                         TRUE ~ FALSE),
    top10_both_bac = case_when(top10_bn == TRUE & top10_dg == TRUE ~ TRUE,
                           TRUE ~ FALSE))

# save hub taxa
write.csv(d.hub_both_bac, row.names = FALSE,
          file = here("Data", "Interim", "Hub_taxa_bac.csv"))

```

### Check how network attributes correlate bacteria
```{r echo=TRUE}
# Check how network attributes correlate
d.hub_both_bac %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~")))

# Highlight OTUs in the top 10% of both_bac degree and betweenness
d.hub_both_bac %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_bac %>% filter(top10_both_bac == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)

(d.hub_both_bac %>% filter(top10_both_bac == TRUE) %>% nrow())/(d.hub_both_bac %>% nrow())


# Highlight OTUs in the top 10 % of either degree or betweenness
d.hub_both_bac %>% 
  ggplot(aes(degree, betweenness)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_bac %>% filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2) 

(d.hub_both_bac %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow())/(d.hub_both_bac %>% nrow())


# Check ranks
d.hub_both_bac %>% 
  ggplot(aes(rank(degree), rank(betweenness))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_regline_equation(aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), 
                                           sep = "~~~~"))) +
  geom_point(data = d.hub_both_bac %>% 
               mutate(degree = rank(degree),
                      betweenness = rank(betweenness)) %>% 
               filter(top10_dg == TRUE | top10_bn == TRUE), aes(degree, betweenness), size = 4, 
             shape = 21, color="red", stroke = 2)


# Number of OTUs that are either or 
d.hub_both_bac %>% filter(top10_dg == TRUE | top10_bn == TRUE) %>% nrow()
d.hub_both_bac %>% filter(top10_dg == TRUE) %>% nrow()
d.hub_both_bac %>% filter(top10_bn == TRUE) %>% nrow()


intersect(d.hub_both_bac %>% filter(top10_dg == TRUE),
          d.hub_both_bac %>% filter(top10_bn == TRUE)) %>% nrow()


d.hub_both_bac %>% filter(top10_both_bac == TRUE) %>% nrow()
d.hub_both_bac %>% filter(top10_bn == TRUE | top10_dg == TRUE) %>% nrow()
```


## Compare different methods
```{r}
## Degree or betweenness in top 10%
top_hub_comb <- d.hub_both %>% filter(top10_bn == TRUE | top10_dg == TRUE ) %>% pull(OTU)
top_hub_fun <- d.hub_both_fun %>% filter(top10_bn == TRUE | top10_dg == TRUE ) %>% pull(OTU)
top_hub_bac <- d.hub_both_bac %>% filter(top10_bn == TRUE | top10_dg == TRUE ) %>% pull(OTU)

# Number of OTUs
top_hub_comb %>% length()
top_hub_fun %>% length()
top_hub_bac %>% length()

# Visualize intersections
## all three
l.otu_top_hub <- list(top_hub_comb = top_hub_comb,
                      top_hub_fun = top_hub_fun,
                      top_hub_bac = top_hub_bac)

UpSetR::upset(UpSetR::fromList(l.otu_top_hub), order.by = "freq")

## Combined kingdom analysis vs. bacteria + fungi
l.otu_top_hub <- list(top_hub_comb = top_hub_comb,
                      top_hub_fun_and_bac = c(top_hub_fun, top_hub_bac))

UpSetR::upset(UpSetR::fromList(l.otu_top_hub), order.by = "freq")

## Check intersect
top_hub_intersect <- intersect(top_hub_comb, c(top_hub_fun, top_hub_bac))

# Check intersection with OTUs of most interest
otu_interest <- c("fOTU860", "fOTU1450", "fOTU1557", "bOTU15", "bOTU5", 
                  "fOTU1598", "bOTU1", "fOTU22", "bOTU4", "fOTU1517", "fOTU1508")

top_hub_intersect <- intersect(top_hub_intersect, otu_interest)
top_hub_intersect

```


```{r}
 
```

